% -*- mode: latex; -*-
% http://en.wikibooks.org/wiki/LaTeX/
\documentclass[10pt]{article}
% \usepackage{abcl}

\begin{document}
\title{An Implementation and Analysis of Addding IRI to Common Lisp's Pathname}
\date{October 2011}
\author{Mark~Evenson}

\maketitle

\section{Abstract}

We implement the semantics for distributed resource description and
retrieval by URL/URI/IRI Pathname in the Armed Bear Common Lisp
implementation.

\section{Plan of Attach}
\subsection{Goals}

\begin{enumerate]
\item Use Common Lisp pathnames to refer to representations referenced
  by a URL.

\item The URL schemes supported shall include at least "http", and
  those enabled by the URLStreamHandler extension mechanism.

\item  Use URL schemes that are understood by the java.net.URL object.


\item  MERGE-PATHNAMES 

        (merge-pathnames "url.asd"
            "http://example/org/armedbear/systems/pgp.asd")
        ==> "http://example/org/armedbear/systems/url.asd"

\item  PROBE-FILE returning the state of URL accesibility.

\item  TRUENAME "aliased" to PROBE-FILE signalling an error if the URL is
not accessible (see "Non-goal 1").

\item  DIRECTORY works for non-wildcards.

\item  URL pathname work as a valid argument for OPEN with :DIRECTION :INPUT.

\item  Enable the loading of ASDF2 systems referenced by a URL pathname.

\item  Pathnames constructed with the "file" scheme
(i.e. #p"file:/this/file") need to be properly URI encoded according
to RFC3986 or otherwise will signal FILE-ERROR.  

\item.  The "file" scheme will continue to be represented by an
"ordinary" Pathname.  Thus, after construction of a URL Pathname with
the "file" scheme, the namestring of the resulting PATHNAME will no
longer contain the "file:" prefix.

\item.  The "jar" scheme will continue to be represented by a jar
Pathname.
\end{enumerate}

\subsection{Non-goals}

\begin{enumerate}

\item We will not implement canonicalization of URL schemas (such as
  following "http" redirects).

\item \textsc{DIRECTORY} will not work for URL pathnames containing
  wildcards.

\end{enumerate}

\subsubsection{Example}
    Example of a Pathname specified by URL:
    
        #p"http://example.org/org/armedbear/systems/pgp.asd"
    


\section{Notes}
\include{notes}

\end{document}

